---
title: "geospatial and regression notebook"
output: html_notebook
---

You've been provided three datasets for this project:

   burglaries_2023.csv: Contains data on the aggravated burglary incidents in Davidson County. This was obtained from https://data.nashville.gov/Police/Metro-Nashville-Police-Department-Incidents/2u6v-ujjs.
   census.csv: Census tract level data on population and median income. This was obtained from the US Census American Community Survey.
   DC: A shapefile containing Davidson County census tracts
Perform a spatial join to determine the census tract in which each burglary occurred. Hint: You may want to make use of the st_as_sf function in order to convert the burglaries data into an sf object.

After performing the spatial join, merge in the census data. Note: Make sure that the final dataset contains all census tracts.

```{r}
library(sf)
library(tidyverse)
library(lubridate)
library(leaflet)
```

```{r}
burglaries <- read_csv("../data/burglaries_2023.csv")
```

```{r}
census <- read_csv("../data/census.csv")
```
Check to see how many rows there are in census?
```{r}
census %>% 
  nrow()
```



```{r}
dc <- st_read("../data/DC/DC.shp")
```
Check to see how many rows there are in dc census tract?
```{r}
dc %>% 
  distinct(TRACTCE) %>% 
  nrow()
```



#convert burglaries into a geospatial object. We can do this using the st_as_sf function.
```{r}
burglaries_converted <- st_as_sf(burglaries |> drop_na(latitude),
         coords = c('longitude', 'latitude'),
         crs = st_crs(dc)
         )
```


#Now, Perform a spatial join to determine the census tract in which each burglary occurred.
```{r}
burglaries_combined <- st_join(dc, burglaries_converted, join = st_contains)
```


#After spatial join, checking to see how many rows there are...
```{r}
burglaries_combined %>% 
  distinct(TRACTCE) %>% 
  nrow()
```

```{r}
census %>% 
  head()
```







```{r}
burglaries_combined<- merge(burglaries_combined, census, by.x = 'TRACTCE', by.y ='tract', all = TRUE)
```

```{r}
burglaries_combined %>% 
  head(2)
```

----
There appears to be the following:
   intplat=latitude
   intptlon=longitude
Q's that you may or may not answer as part of your EDA... 
   1. why does the index follow an odd numbering schema?
   2. Does countyfp or statefp ever change?
   3. What is the distribution of times and days for incidents (need to convert date time...tried, but still have weird seconds thing...)
   4. Various info on offense descriptions, weapons, victims, 
----
Part 2 - Exploratory Analysis
Perform some exploratory analysis on your prepared dataset.

Aggregate the data by census tract. Warning: each incident can appear multiple times if there are multiple victims, so be sure that you aren't double-counting any incidents.

Which census tract had the highest number of burglaries? Which census tract had the highest number of burglaries per 1000 residents?

We're interested in the relationship between median income and number of aggravated burglaries, so examine those variables on their own and together to see what you can find. You may want to perform additional calculations, create plots, etc.

Initial EDA

```{r}
summary(burglaries_combined)
```

```{r}
burglaries_combined %>% 
  head(2)
```
```{r}
burglaries_combined %>% 
  tail(2)
```
We see here that there are 174 census tracts and they are connected to geometry...
```{r}
census_by_tract<-burglaries_combined |> 
  count(TRACTCE)
census_by_tract
```
Note a few things about the answer above
  "NAD 83 provides a frame of reference for latitude and longitude locations on Earth. Surveyors now rely almost exclusively on the Global Positioning System (GPS) to identify locations on the Earth and incorporate them into existing geodetic datums. For example, NAD27, NAD83, and WGS84 are the most common geodetic datums in North America."

Count by time.  (How does time work in R? Do we have to do dt calcs?)
```{r}
count_by_time<-burglaries_combined |> 
  count(incident_occurred)
count_by_time
```
```{r}
burglaries_combined <- burglaries_combined |> 
  mutate(incident_difference = incident_reported - incident_occurred)
burglaries_combined
```


Why doesn't this work?
```{r}
library(lubridate)
burglaries_combined %>% 
 seconds_to_period(incident_difference) 

```

```{r}

burglaries_combined <- burglaries_combined %>% 
 mutate(incident_difference_hrs = incident_difference/(60*60)) 
burglaries_combined
```

```{r}
burglaries_combined |> 
  st_drop_geometry() |> 
  group_by(weapon_description) |> 
  count(name = "weapons") |> 
  arrange(desc(weapons))
```

```{r}
burglaries_combined |> 
  st_drop_geometry() |> 
  group_by(victim_type) |> 
  count(name = "victim") |> 
  arrange(desc(victim))
```

```{r}
burglaries_combined |> 
  st_drop_geometry() |> 
  group_by(victim_number) |> 
  count(name = "victim") |> 
  arrange(desc(victim))
```

```{r}
burglaries_combined |> 
  ggplot() +
  geom_sf()
```

```{r}
burglaries_combined |> 
  ggplot() +
  geom_sf(aes(fill = zip_code))
```

```{r}
ggplot(burglaries_combined, aes(x=zip_code)) +
  geom_histogram(bin=30, color='black', fill='blue') 
```


Aggregate the data by census tract. Warning: each incident can appear multiple times if there are multiple victims, so be sure that you aren't double-counting any incidents.

  #Use group-by nunique?(use distinct, also it must go after groupby)
  #filter by unique incident_number, then aggregate by TRACTCE

Initial Aggregation: (before aggregating out individual incidents)
  #note that some drop the geometry, using   st_drop_geometry() |>   not included here.  
```{r}
burglaries_combined |> 
  group_by(TRACTCE) |> 
  count(name = "count") |> 
  arrange(desc(count))

```
Secondary Aggregation: (after aggregating out individual incidents)
Burglaries aggregated by incident number...
#note that some drop the geometry, using   st_drop_geometry() |>   not included here.  
(You could take this and then make another one that subsets out the population and knit it together...)
```{r}
burglaries_count_per <- burglaries_combined |> 
  group_by(TRACTCE) |> 
  distinct(incident_number) |> 
  count(name = "incident_number") |> 
  arrange(desc(incident_number))
burglaries_count_per
```
Incident_number counts
```{r}
incidents_agg_count <- burglaries_combined |> 
  group_by(incident_number) |> 
  count(name = "count") |>
  arrange(desc(count))
incidents_agg_count
```

looking at the count of incident numbers graphically.  
```{r}
ggplot(incidents_agg_count, aes(x=count)) +
  geom_bar( color='black', fill='blue') 

```


Which census tract had the highest number of burglaries?  
(016000)

Which census tract had the highest number of burglaries per 1000 residents? (016000) (unless you count the two tracts with no population, then 980100...)
  #to get this number we have to subset number of incidents by population.  Perhaps this is best done with a separate column...


```{r}
burglaries_pop <- burglaries_combined |>
  select(TRACTCE, population)|>
  group_by(TRACTCE) |> 
  distinct(population) 
```

```{r}
burglaries_combined |> 
  group_by(TRACTCE) |> 
  distinct(incident_number) |> 
  count(name = "incident_number") |> 
  arrange(desc(incident_number))
```

```{r} 
burglaries_joined <- merge(burglaries_count_per, burglaries_pop)

```

```{r}
burglaries_joined <- burglaries_joined %>% 
  mutate(burglaries_per_capita = population/incident_number) %>% 
  arrange(burglaries_per_capita)
burglaries_joined
```


We're interested in the relationship between median income and number of aggravated burglaries, so examine those variables on their own and together to see what you can find. You may want to perform additional calculations, create plots, etc

```{r}
ggplot(burglaries_joined, aes(x=population, y=burglaries_per_capita)) +
  geom_point(color='darkblue', fill='black') 

```
```{r}
burglaries_median <- burglaries_combined |>
  select(TRACTCE, median_income, geometry) |>
  group_by(TRACTCE) |> 
  distinct(median_income) 
burglaries_median
```
```{r} 
burglaries_joined <- merge(burglaries_joined, burglaries_median)

burglaries_joined %>% 
  arrange(desc(incident_number))
```


(need to subset some of the data here.  There are several observations that are -666666666 in median income...)
```{r}
burglaries_joined<- burglaries_joined %>% 
  filter(median_income > 0) %>% 
  arrange(desc(median_income))
```



```{r}

ggplot(burglaries_joined, aes( x=incident_number, y=median_income, hue=population)) +
  geom_point(color='darkblue') 

```

(Map with burglaries. Darker is more incidents.)

```{r}
incidents_agg_count %>% 
  ggplot() +
  geom_sf(aes(fill = incident_number))
```

Part 3 - Statistical Modeling
Fit a Poisson regression model with target variable the rate of burglaries per census tract and with predictor the median income. Offset using the log of the population so that we are looking at the rate of burglaries per population instead of the number of burglaries. How can you interpret the meaning of the output? How do the estimates from the model compare to the observed data?

target variable the rate of burglaries per census tract 
   #I assume this means we are using time?
   #In the example the 'rate' was just the count of the visits
   #Ran here without a time component, just using number of unique incidents per tract

```{r}
burglaries_joined |> 
  ggplot(aes(x = incident_number)) + 
  geom_bar(color='black', fill='blue')
```

```{r}

pois_num_0 <- glm('incident_number ~ 1',
               data = burglaries_joined,
               family = poisson)
summary(pois_num_0)

```

```{r}

pois_num_1 <- glm('incident_number ~ median_income',
               data = burglaries_joined,
               family = poisson)
summary(pois_num_1)

```

Coefficients
(To get the estimates for something in a poisson, you have to exponentiate the results to get the average)
For a one unit change in median_income, the difference in the logs of expected counts of burglaries is expected to change by -1.958e-05
Or, for every dollar increase in median income, burglaries are expected to decrease by 0.00001958 in logs of expected counts.
Oftentimes the poisson results are converted to 'incident response ratios' (IRR) this is done by taking the exponential of the result
or exp(-1.958e-05) = .99998402.  This represents the multiplicative change in the expected count for a one-unit increase in the predictor.
So, another way to interpret this is to say that for every one-unit increase in median income, the IRR decreases by 0.99998042, or a little less than one percent. 
Finally, For every one-unit increase in "median_income," the expected count of the response variable is multiplied by approximately
Further, AIC. Stands for Akaike information criterion.  Basically, the lower the AIC, the better the model fit. Compare to the other 3 pois_reg for fit comparisons...

Let's see what it estimates for the mean number of the distribution:
```{r}
mu = exp(coef(pois_num_1 ))
```


And plot the result. (these are means...)
```{r}
x <- 0:39
y <- dpois(x, lambda = mu)

tibble(x = x, y = y) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_col(color='black', fill='green')
```
So just so we're clear here, this means there are 40 tracts with 1 incident report, and 24 tracts with 2...1 tract with 39. And, in the above graph we're taking the means of all of those as a percentage of the total.
```{r}
burglaries_joined|> 
  count(incident_number)
```

Offset using the log of the population so that we are looking at the rate of burglaries per population instead of the number of burglaries

```{r}
burglaries_joined <- burglaries_joined |> 
  mutate(poplog = log10(population))

```


```{r}
pois_num_2 <- glm('incident_number ~ median_income',
               data = burglaries_joined,
               family = poisson,
               offset = log(population))
summary(pois_num_2)
```

For every one unit increase of median_income, the IRR decreases by 

```{r}
pois_num_3 <- glm('incident_number ~ median_income',
               data = burglaries_joined,
               family = poisson,
               offset = poplog)
summary(pois_num_3)

```

```{r}
burglaries_joined |> 
  mutate(group = ntile(median_income, n = 10)) |> 
  group_by(group) |> 
  summarize(
    median_income = median(median_income),
    median_incident_number = median(incident_number)
    )
```

```{r}
burglaries_joined |> 
  mutate(group = ntile(median_income, n = 10)) |> 
  group_by(group) |> 
  ggplot(aes(x = incident_number)) +
  geom_bar(color='black', fill='darkred') +
  facet_wrap(~group)
```

```{r}

burglaries_joined<- burglaries_joined |> 
  mutate(group = ntile(median_income, n = 5))
```


```{r}
mu3 = exp(coef(pois_num_3 ))
```


And plot the result. (these are means...)
```{r}
x <- 0:35
y <- dpois(x, lambda = mu3)

tibble(x = x, y = y) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_col(color='black', fill='orange')
```

```{r}

ggplot(burglaries_joined, aes( x=incident_number, y=population, hue=group)) +
  geom_point(color='darkblue') 
```



```{r}
est_df <- tibble(
  median_income = seq(from = min(burglaries_joined$median_income, na.rm = TRUE), 
                          to = max(burglaries_joined$median_income, na.rm = TRUE), 
                          length.out = 100),
  population = seq(from = min(burglaries_joined$population, na.rm = TRUE), 
                          to = max(burglaries_joined$population, na.rm = TRUE), 
                          length.out = 100),    )

est_df <- est_df %>%  
  bind_cols(predict(pois_num_2, newdata = est_df, type = "response")) %>% 
  rename("Estimated Mean Number of Incidents" = "...3")

ggplot(est_df) +
  geom_line(aes(x = median_income, y = `Estimated Mean Number of Incidents`)) +
  geom_point(data = burglaries_joined, aes(x=median_income, y=incident_number))
           
```
```{r}
est_df %>%  
  ggplot(aes(x = median_income, y = `Estimated Mean Number of Incidents`)) +
  geom_line(color = 'green')
```

