---
title: "geospatial and regression notebook"
output: html_notebook
---

You've been provided three datasets for this project:

   burglaries_2023.csv: Contains data on the aggravated burglary incidents in Davidson County. This was obtained from https://data.nashville.gov/Police/Metro-Nashville-Police-Department-Incidents/2u6v-ujjs.
   census.csv: Census tract level data on population and median income. This was obtained from the US Census American Community Survey.
   DC: A shapefile containing Davidson County census tracts
Perform a spatial join to determine the census tract in which each burglary occurred. Hint: You may want to make use of the st_as_sf function in order to convert the burglaries data into an sf object.

After performing the spatial join, merge in the census data. Note: Make sure that the final dataset contains all census tracts.

```{r}
library(sf)
library(tidyverse)
```

```{r}
burglaries <- read_csv("../data/burglaries_2023.csv")
```

```{r}
census <- read_csv("../data/census.csv")
```
Check to see how many rows there are in census tract?
```{r}
census %>% 
  nrow()
```



```{r}
dc <- st_read("../data/DC/DC.shp")
```

#convert burglaries into a geospatial object. We can do this using the st_as_sf function.
```{r}
burglaries_converted <- st_as_sf(burglaries |> drop_na(latitude),
         coords = c('longitude', 'latitude'),
         crs = st_crs(census)
         )
```


Now, Perform a spatial join to determine the census tract in which each burglary occurred.
```{r}
burglaries_combined <- st_join(census, burglaries_converted, join = st_contains)
```


After spatial join, checking to see how many rows there are...
```{r}
burglaries_combined %>% 
  distinct(TRACTCE) %>% 
  nrow()
```
```{r}
summary(burglaries_combined)
```







